---
title: "Hypothesis Testing - Homework 5"
output: html_notebook
---
Consider the in-built R dataset "sleep". This dataset compares the effects of two sleep medications. That is, the increase in number of hours of sleep in patients in comparison to a control.
```{r}
sleep
```
```{r}
extra <- sleep[,1]
group <- c(1,1,1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,2,2)
boxplot(extra ~ group)
```
We immediately see the group one drug produces, on average, less of an increase in hours of sleep for each patient. This gives us a feel for whether we expect there to be a statistical difference. In order to test whether there is a difference between the two drugs effects we wish to test $H_0: \mu_1 - \mu_2 = 0$ against $H_1: \mu_1 - \mu_2 \ne 0$ where $\mu_1$ and $\mu_2$ are the mean extra hours slept for group one and group two respectively.

We first check whether the data are normally distributed. We check this by performing a qqplot of the patients in each group and checking the data is roughly normally distributed.
```{r}
extra1 <- extra[1:10]
extra2 <- extra[11:20]

qqnorm(extra1)
qqline(extra1)
qqnorm(extra2)
qqline(extra2)
```

Hence we see they are suitably normally distributed and so we wish to introduce a model using pairs of random variables $(X_i, Y_i)$. Let $D_i = X_i-Y_i$ for all $i$ then $D \sim N(\mu, \sigma^2)$ with unknown parameters $\mu$ and $\sigma^2$. We therefore use the t-test, test statistic. That is,

$$\frac{\sqrt{n} ( \bar{D}-\mu_0)}{S_D} \sim t(n-1)$$

under the null hypothesis $H_0: \mu = \mu_0$ where $S_D^2$ is the sample variance of the differences.
```{r}
n <- 10 #there are ten patients
difference <- extra1 - extra2
SD <- sqrt(var(difference))

tstat <- (sqrt(n)*(mean(difference) - 0))/SD
tstat
```
We know this is then distributed according to a $t(9)$ distribution so we calculate the critical value at the $\alpha = 0.05$ level.
```{r}
critval <- qt(0.95, 9)
critval
```
Since we see the t-statistic exceeds the critical value, at the 0.95 confidence level wecan reject the nnull hypothesis and conclude there is a statistical difference in number of hours slept between patients on the two medications.

We compare our calcualtions to R's built-in t.test function
```{r}
t.test(difference, alternative="two.sided")
```

