---
title: "MLE of The Exponential Distribution - Homework 4"
output: html_notebook
---
We wish to estimate the parameter $\lambda$ of an exponential distribution using the method of maximum likelihood estimation. This method involves finding parameter values which maximise the likelihood that the process described by the model produced the observed data.

Firstly we need to derive the log-likelihood function for the exponential distribution. Suppose $X \sim \text{exponential}(\lambda)$ then the joint pdf of the random variables $X_1, X_2, \cdots, X_n$ is given by $$\begin{align*} f_\textbf{x}(\textbf{x}) &= \Pi_{i=1}^nf(x_i) \\ &= \Pi_{i=1}^n \lambda e^{-\lambda x} \\ &= \lambda^n \Pi_{i=1}^n e^{-\lambda x} \\ &= \lambda^n e^{-\lambda \Sigma_{i=1}^n x_i} \\ &= \lambda^n e^{-\lambda n \bar{x}}. \end{align*}$$

We then find the log-likelihood function as the logarithm of the above function, that is,

$$\begin{align*} l(\lambda, \textbf{x}) &= \log f_\textbf{x}(\textbf{x}) \\ &= \log(\lambda^n e^{-\lambda n \bar{x}}) \\ &= \log(\lambda^n)+\log(e^{-\lambda n \bar{x}}) \\ &= n\log(\lambda) - \lambda n \bar{x}. \end{align*}$$

We then add this function into R
```{r}
exp_log_lik <- function(y, lambda) {
  n <- length(y)
  logl <- n*log(lambda) - lambda*n*mean(y)
  return(-logl)
}
```

Next we produce a function to generate a random sample of 10000 data points from an $\text{exponential}(\lambda)$ distribution. We will use the 'optim' function to generate the MLE estimate of $\lambda.$ We finally compare the values of the expectation for the exponential distribution to the MLE estimate for a range of $\lambda.$ These data are tabulated below. (Note a NaN error is thrown here, and I am unable to pinpoint its source, however it has no visible effect on the outcome of the code.)

```{r}
mle_exp <-function(lambda) {
  exp_data <- rexp(10000, lambda) #Generate sample from expon: n=10000, lambda=lambda
  
  direct_lambda_est <- 1/mean(exp_data) #Generate a lambda est directly
  res_exp <- optim(par=lambda, exp_log_lik, y=exp_data, method="BFGS")

  estimate_list <- list(direct_lambda_est, res_exp$par) #Must use list to access
  return(estimate_list) #Return MLE and direct estimates
}

#Define vectors containing estimates and lambda values
lambdas <- seq(1,10)
direct_estimates <- c()
mle_estimates <- c()

#Loop through each lambda value and calculate estimates
for (lambda in lambdas) {
  exp_sample <- mle_exp(lambda) #Calculate here for same sample for both estimates

  direct_estimates <- c(direct_estimates, exp_sample[1])
  mle_estimates <- c(mle_estimates, exp_sample[2])
}

#Tabulate and display the data
estimate_frame = data.frame(Lambda = lambdas, Direct = unlist(direct_estimates), MLE = unlist(mle_estimates))
estimate_frame
```
We see from this the MLE estimate is an excellent approximation to the parameter. It becomes better with increasing sample size $n$.
