---
title: "MATH2715 - Homework 2 - Normal Distrubtion"
output: html_notebook
---
Let $X \sim N(\mu, \sigma^2).$ We define a function 'norm' that returns $\Pr(a < X \leq b) \text{ for all } a \leq b.$ It makes use of the inbuilt R function 'pnorm', which takes parameter $\alpha$ returns the value of the integral $\int_{-\infty}^{\alpha}\frac{1}{\sigma\sqrt{2\pi}} \exp\Big(-\frac{(x-\mu)^2}{2\sigma^2}\Big).$ We specifically evaluate 'norm' for $a=-2$ and $b=3$ where $X \sim N(1, 2^2).$
```{r}
norm <- function(a, b, mu, sigma) {
  #pnorm(x) returns the integral of the pdf from -inf to x.
  p <- pnorm(b, mean=mu, sd=sigma) - pnorm(a, mean=mu, sd=sigma)
  return(p)
}

norm(-2, 3, 1, 2)
```
Nextly, we plot the pdf and cdf of $X \sim N(1, 2^2).$ We do this by generating a sequence of values with a suitable start and endpoint. The sequence should also increment in steps of 0.1, this is adequately small to ensure the curve is smooth.

For the sequence, 'x', we then calculate a vector 'freq' which contains the value of the pdf of the normal distribution for each element in the sequence 'x'. Plot 'freq' against 'x' to see the distribution.

For the sequence, 'x', calculate a new vector 'cfreq' which contains the value of the cdf of the normal distribution for each element in the sequence 'x'. Plot 'cfreq' against 'x' to see the cumulative distribution.
```{r}
x <- seq(-6, 8, by = 0.1) #increment in steps of 0.1 for a smooth curve.
freq <- dnorm(x, mean=1, sd=2)
plot(x, freq, type = "l" , ylab="Density", lwd=2, main="pdf of N(1, 2^2)")

cfreq <- pnorm(x, mean=1, sd=2)
plot(x, cfreq, type = "l" , ylab="Probability Density", lwd=2, main="cdf of N(1, 2^2)")
```
A Q-Q plot is a plot of the theoretical quantiles against empirical (expermiental) quantiles of a given distribution. We wish to show the Q-Q plot of $N(1, 2^2)$. We use R's inbuilt function 'qnorm' to calculate the theoretical quantiles of the distribution for 0.00, 0.05, 0.10, ..., 1.00. We use the inbuilt function 'rnorm' to generate a random sample of size 100 from $N(1, 2^2)$. Plot these against each other to see the Q-Q plot. In general, if a model is a good fit, each emprirical quantile will match its respective theoretical quantile and the datapoints will lie close to the line y=x. I have added this line for comparison.
```{r}
quantilesample <- seq(0,1,0.01) #sample every quantile.
theoreticquantiles <- qnorm(quantilesample, mean=1, sd=2)
empiricalquantiles <- quantile(rnorm(100, mean=1, sd=2), probs = quantilesample)
empiricalquantiles <- as.numeric(empiricalquantiles)

plot(theoreticquantiles, empiricalquantiles, main="Q-Q Plot of N(1,2^2)")
abline(0,1) #add the line y=x
```

